The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2019-03-08 Paper.Rmd
Discussion : 1st para : summary added

<!--what we did and why is it important to study?-->
<!--***DAN: some notes by which to revise this para, Shya:
1) no need to restate methods at all, the methods are not among our main results in this paper
2) No real need to say much about why it is important to study what we stdied, that was done in the intro. Except insofar as saying what our main results are should make it clear it is an important thng to study, and see also further notes below.
3) Some additional points below
-->
<!--
Most studies in ecology have focused on measuring correlations between variables and thus ignoring the possibility to recognize importance of stronger association between extreme values. In the present study, we introduced the concept of asymmetric tail-dependence and non-Normal copulae in this field. -->
<!--***DAN: Too discursive. We do not need to repeat that we introduced this. Just summarize our main results in pothy language! And then maybe emphasize the importance, or make any additional points about importance that could not be made in the Intro. This para needs to be hard-hitting and brief, saying all the important thigs we accomplished in short form, all in one place. See below for some examples.-->

@@

<!--***DAN: How about simply: "We showed that non-N copula structure and asymmetric tail dependence are exceedingly common in datasets that range across multiple sub-disciplines of ecology." See if you can write 1-3 pithy, impactful sentences like this one for Q2, and 1-3 for Q3 as well. Look at the first para of Discussion for some of my recent papers to get examples of the writing techniques for how we have done this in the past.-->
<!--We applied two methods. First, we did a model selection approach based on maximum likelihood which fit the real copula data with an optimal set of 16 copulae family. Copula having highest AIC-weight came as best-fit from this approach. Though we did goodness of fit test based on two statistics (Cramer-von Mises and Kolmogorov-Smirnov), we wanted to ensure the reliability of our model selection results. So, on a second note, we measured the tail dependence at both lower and upper end of a copula by three different nonparametric statistics ($\cor$, $\Ps$ and $\Dsq$) which we defined and we found they were consistent with the inferences we drew from model selection results. -->
<!--***DAN: No need at all to summarize methods again, except in cases where the methods are 
among the main products of the paper - not our case. So delete the above text. If you want to discuss our methods, do in later in the Discussion, not in this first paragraph, unless you may want to have a sentence saying something like "We synthesized and expanded upon methods for statistically assessing copula structure and tail dependence, and models for exploring the causes and consequences of copula strucuture in ecology and related fields."-->
<!--Our approach could help to interpret the synchronous behavior of a species from several meta-populations. To illustrate, if pairwise copulae from a multivariate spatial data of a species (e.g. abundance data of green-spruce aphid) showed major "lower-tail dependence" on most of the locations then it could be considered as a signature of "spatial synchrony" that meta-populations are simultaneously rare at those locations. Conversely, if some aphid species showed synchronously "upper-tail dependence" then it would be harmful for corp-production or in case of synchronous abundance of plankton could lead to algal-bloom in the ocean. Copulae could also give us information about the causal relationship between ecological variables. For example, if maximum rainfall is beneficial for some species then it would show an upper-tail dependent copula where high values of species abundance and precipitation would be more strongly correlated than lower values. -->
<!--***DAN: I do not see how these examples go beyond the examples of the introduction. If they do, please clarify. If they don't, please remove.-->

>>>>

We showed that non-N copula structure and asymmetric tail dependence are 
exceedingly common in datasets that range across multiple sub-disciplines
of ecology. Via basic mathematical models, we hypothesized two probable
mechanisms causing those non-N copula structures with asymmetric tail-dependence. One
mechanism was "Moran"-like effect where dependence structure between ecological
variables from two locations was shaped by the dependence structure of 
the environmental variables at those patches. Second mechanism, a mimic of 
Liebig's law of the minimum, explored
the asymmetric influence of environmental variables which 
could limit the growth of populations when below (or beyond)
a certain threshold. Using model simulations and real ecological data,
we concluded asymmetric tail dependence feature of 
non-N copula could affect (1) the skewness (henceforth, the stability)
of spatially averaged quantities over years, (2) the extinction risk 
and (3) the spatial Taylor's law slope of metapopulations.

<!--***DAN: some notes by which to revise this para, Shya:
1) no need to restate methods at all, the methods are not among our main results in this paper
2) No real need to say much about why it is important to study what we stdied, that was done in the intro. Except insofar as saying what our main results are should make it clear it is an important thng to study, and see also further notes below.
3) Some additional points below
-->
<!--
Most studies in ecology have focused on measuring correlations between variables and thus ignoring the possibility to recognize importance of stronger association between extreme values. In the present study, we introduced the concept of asymmetric tail-dependence and non-Normal copulae in this field. -->
<!--***DAN: Too discursive. We do not need to repeat that we introduced this. Just summarize our main results in pothy language! And then maybe emphasize the importance, or make any additional points about importance that could not be made in the Intro. This para needs to be hard-hitting and brief, saying all the important thigs we accomplished in short form, all in one place. See below for some examples.-->

@@

<!--***DAN: How about simply: "We showed that non-N copula structure and asymmetric tail dependence are exceedingly common in datasets that range across multiple sub-disciplines of ecology." See if you can write 1-3 pithy, impactful sentences like this one for Q2, and 1-3 for Q3 as well. Look at the first para of Discussion for some of my recent papers to get examples of the writing techniques for how we have done this in the past.-->

<!--We applied two methods. First, we did a model selection approach based on maximum likelihood which fit the real copula data with an optimal set of 16 copulae family. Copula having highest AIC-weight came as best-fit from this approach. Though we did goodness of fit test based on two statistics (Cramer-von Mises and Kolmogorov-Smirnov), we wanted to ensure the reliability of our model selection results. So, on a second note, we measured the tail dependence at both lower and upper end of a copula by three different nonparametric statistics ($\cor$, $\Ps$ and $\Dsq$) which we defined and we found they were consistent with the inferences we drew from model selection results. -->
<!--***DAN: No need at all to summarize methods again, except in cases where the methods are 
among the main products of the paper - not our case. So delete the above text. If you want to discuss our methods, do in later in the Discussion, not in this first paragraph, unless you may want to have a sentence saying something like "We synthesized and expanded upon methods for statistically assessing copula structure and tail dependence, and models for exploring the causes and consequences of copula strucuture in ecology and related fields."-->
<!--Our approach could help to interpret the synchronous behavior of a species from several meta-populations. To illustrate, if pairwise copulae from a multivariate spatial data of a species (e.g. abundance data of green-spruce aphid) showed major "lower-tail dependence" on most of the locations then it could be considered as a signature of "spatial synchrony" that meta-populations are simultaneously rare at those locations. Conversely, if some aphid species showed synchronously "upper-tail dependence" then it would be harmful for corp-production or in case of synchronous abundance of plankton could lead to algal-bloom in the ocean. Copulae could also give us information about the causal relationship between ecological variables. For example, if maximum rainfall is beneficial for some species then it would show an upper-tail dependent copula where high values of species abundance and precipitation would be more strongly correlated than lower values. -->
<!--***DAN: I do not see how these examples go beyond the examples of the introduction. If they do, please clarify. If they don't, please remove.-->
